\vspace{-3em} \textbf{ Mapping stellar surfaces\\ II: An interpretable Gaussian process model for light curves }

### 2021-01-27 paper2/ms.tex
paper1 bugfix
override minor issue with kwargs in one of the runs

% Title
\title{%
    \vspace{-2em}
    \textbf{
        Mapping stellar surfaces\\
        II: An interpretable Gaussian process model for light curves
    }
}

@@

Over the past decades, Gaussian processes (GPs) have gained traction as
a leading tool for modeling correlated signals in astronomical time series datasets, especially photometric time series
\citep[e.g.,][]{RasmussenWilliams2005,BrewerStello2009,Luger2016,ForemanMackey2017,Angus2018} and radial velocities (ADD CITATIONS).
\dfm{LET'S START WITH A SIMPLE INTUITIVE INTRO HERE AND SAVE THE DETAILS UNTIL LATER! HERE'S A VERY ROUGH FIRST PASS:
GPs are popular models for these applications because they allow marginalization over a stochastic noise process specified only by its autocorrelation structure or kernel.
This is relatively straightforward to interpret and in some specific cases it can be related to physically motivated models (CITE SOME THINGS).
Furthermore, there are several popular open source implementations that allow efficient evaluation of these models, and these have been widely demonstrated to be useful effective models for the time series when the stochastic variability is primarily a nuisance.}

% However, an important drawback of the typical uses of GPs when applied
% to modeling stellar light curves, particularly when the variability is due
% to starspots rotating in and out of view.
If, however, one hopes to make inferences about the stellar surface itself, GP models, as they are typically applied, are not so useful.
\textbf{While it is straightforward to derive
    posterior constraints on the hyperparameters of an effective GP model for observations of a star, it is not clear what those constraints actually tell us about the stellar surface.}
Specifically, in all but a few restricted cases, there is no first principles relationship between the descriptive parameters of a typical GP model (see section WHATEVER below)
and the physical properties of the stellar surface surface that is being observed.
For instance, it may be tempting to interpret the amplitude of the variability as some measure of the spot
contrast or the total number of spots, or the timescale of the variations as the
spot lifetime, but there are no guarantees these interpretations will
hold in general. After all, the choice of kernel is quite often \emph{ad hoc},
providing an \emph{effective}---as opposed to \emph{interpretable}---description
of the physics.
%
There are two important exceptions to this: asteroseismic studies, in which the
the GP hyperparameters can offer direct insight into
the behavior of complex pulsation modes and thus physical properties of the
stellar interior \citep[e.g.,][]{BrewerStello2009,ForemanMackey2017}; and
stellar rotation period studies, in which the period hyperparameter $P$ can usually
be associated with the rotation period of the star \citep[e.g.,][]{Angus2018}.%
\footnote{An exception to this is in the presence of strong differential
    rotation, in which case many periods may be present in the data, or
    when spots evolve coherently, which can also introduce weak periodicities
    in the light curve.}
%
For spot-induced variability, on the other hand,
GPs are usually used
when the variability itself is a nuisance parameter. For example, if the
goal is to constrain the properties of a transiting exoplanet or to
search for a planetary signal in a radial velocity dataset, a GP might be
used to remove (or, better yet, to marginalize over) the stellar variability
\citep[e.g.,][]{Haywood2014,Rajpaul2015,Luger2017b}.
In this case, the physics behind the variability is irrelevant, so an
effective model of this sort is generally sufficient.

>>>>

% Title
\title{%
    \vspace{-3em}
    \textbf{
        Mapping stellar surfaces\\
        II: An interpretable Gaussian process model for light curves
    }
}

@@

Over the past decades, Gaussian processes (GPs) have gained traction as
a leading tool for modeling correlated signals in astronomical time series datasets,
especially photometric time series
\citep[e.g.,][]{RasmussenWilliams2005,BrewerStello2009,Luger2016,ForemanMackey2017,Angus2018}
and radial velocities (ADD CITATIONS).
GPs are popular models for these applications because they allow marginalization
over a stochastic noise process specified only by its autocorrelation structure
or kernel.
This is relatively straightforward to interpret and in some specific cases it
can be related to physically motivated models (CITE SOME THINGS).
Furthermore, there are several popular open source implementations that allow
efficient evaluation of these models, and these have been widely demonstrated
to be useful effective models for the time series when the stochastic
variability is primarily a nuisance.

% However, an important drawback of the typical uses of GPs when applied
% to modeling stellar light curves, particularly when the variability is due
% to starspots rotating in and out of view.

If, however, one hopes to make inferences about the stellar surface itself,
GP models, as they are typically applied, are not so useful.
\textbf{While it is straightforward to derive
    posterior constraints on the hyperparameters of an effective GP model for
    observations of a star, it is not clear what those constraints actually
    tell us about the stellar surface.}
Specifically, in all but a few restricted cases, there is no first principles
relationship between the descriptive parameters of a typical GP model
(see section WHATEVER below)
and the physical properties of the stellar surface surface that is being observed.
For instance, it may be tempting to interpret the amplitude of the variability
as some measure of the spot contrast or the total number of spots,
or the timescale of the variations as the
spot lifetime, but there are no guarantees these interpretations will
hold in general. After all, the choice of kernel is quite often \emph{ad hoc},
providing an \emph{effective}---as opposed to \emph{interpretable}---description
of the physics.
%
There are two important exceptions to this: asteroseismic studies, in which the
the GP hyperparameters can offer direct insight into
the behavior of complex pulsation modes and thus physical properties of the
stellar interior \citep[e.g.,][]{BrewerStello2009,ForemanMackey2017}; and
stellar rotation period studies, in which the period hyperparameter $P$ can usually
be associated with the rotation period of the star \citep[e.g.,][]{Angus2018}.%
\footnote{An exception to this is in the presence of strong differential
    rotation, in which case many periods may be present in the data, or
    when spots evolve coherently, which can also introduce weak periodicities
    in the light curve.}
%
For spot-induced variability, on the other hand,
GPs are usually used
when the variability itself is a nuisance parameter. For example, if the
goal is to constrain the properties of a transiting exoplanet or to
search for a planetary signal in a radial velocity dataset, a GP might be
used to remove (or, better yet, to marginalize over) the stellar variability
\citep[e.g.,][]{Haywood2014,Rajpaul2015,Luger2017b}.
In this case, the physics behind the variability is irrelevant, so an
effective model of this sort is generally sufficient.
